---
title: 'Milestone 1 '
author: 'Racquelle & Farnaz '
date: "27/02/2020"
output:
  html_document: default
  pdf_document: default
---

```{r fake-load-libraries, eval=FALSE}
library(tidyverse)
library(magrittr)
library(here)
```

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

# Task 1: Choosing a dataset

We found the dataset at: https://www.kaggle.com/russellyates88/suicide-rates-overview-1985-to-2016/data

# Task 2: Project Proposal & EDA

#### 2.1: Introduce and describe your dataset

This dataset 

#### 2.2: Load your dataset

```{r}
suiciderates<- read.table(("suiciderates.csv"),sep=" ")
```

Peek at dataset:
```{r}
DT::datatable(suiciderates)
```

#### 2.3: Explore your dataset

> Exploratory Data Analysis of 'suiciderates'

How many rows?
```{r}
nrow(suiciderates)
```

How many columns?
```{r}
ncol(suiciderates)
```

Summary of suiciderates dataset:
```{r}
summary(suiciderates)
```

> Figuring out NAs in 'suiciderates' dataset:

Out of entire dataset (27820 observations of 12 variables), what % are NAs?
```{r}
sum(is.na(suiciderates))/27820*12
```

For column Human Development Index (HDI) for year, what % are NAs?
```{r}
sum(is.na(suiciderates$HDI.for.year))/27820
```
Since there are 8.39% of NAs in the dataset, and the variable 'HDI for year' consists of 70% NAs, we have decided to drop that variable in our analyses, since 'HDI for year' values wouldn't be significant to factor in when looking at suicide rates due to lack of data.

> Removing NAs and creating refined 'suicideratesnew' dataset:

Next, I will select for only the variables I am interested in, thus removing 'HDI for year'. 

```{r}
suicideratesnew <- suiciderates %>% 
  select(country, year, sex, age, suicides_no, population, suicides.100k.pop, country.year, 
         gdp_for_year...., gdp_per_capita....,generation)
```

```{r}
DT::datatable(suicideratesnew)
```

I will now check to see how many NAs are still remaining in this dataset:
```{r}
sum(is.na(suicideratesnew))/27820*11
```
There are now 0% of NAs in the new dataset, further exemplifying that 'HDI for year' contained all the NAs.

> Exploratory Data Analysis of 'suicideratesnew':

How many rows?
```{r}
nrow(suicideratesnew)
```

How many columns?
```{r}
ncol(suicideratesnew)
```

Summary of suiciderates dataset:
```{r}
summary(suicideratesnew)
```